---
title: "Education Surveys"
output: html_notebook
---

# Data
```{r}
# Load necessary libraries
library(dplyr)
library(tidyr)
library(ez)
library(writexl)
library(ggplot2)
library(emmeans)
library(car)
library(WRS2)
library(lme4)
library(patchwork)
library(viridis)
library(cowplot)
library(ggpubr)
library(gt)

# Load the data (replace 'survey_data.csv' with your actual file path)
data <- read.csv('/Users/alexgood/Desktop/Education Chapter/Survey_stats_new.csv')
View(data)

# Count unique student IDs
unique_ids <- data %>%
  summarise(num_unique_ids = n_distinct(`ID`))
# Print the result
print(unique_ids) # 100 responses 

table(unique(data[c("ID", "Gender")])$Gender)

# Convert relevant columns to appropriate data types
data$Survey <- as.factor(data$Survey)
data$ID <- as.factor(data$ID)
data$Teacher <- as.factor(data$Teacher)
data$Grade <- as.factor(data$Grade)
data$ID <- as.factor(data$ID)
data$Course <- as.factor(data$Course)
data$Period <- as.factor(data$Period)
data$Environment <- as.numeric(data$Environment)
data$Corals <- as.numeric(data$Corals)
data$Last <- as.factor(data$Last)
data$First <- as.factor(data$First)
data$Gender <- as.factor(data$Gender)

summary(data)

# Reshape the data into long format (if not already in long format)
long_data <- data %>%
  pivot_longer(
    cols = c(Attitude, Career, Environment, Corals, Total),
    names_to = "Category",
    values_to = "Score"
  )

long_data$ID <- as.factor(long_data$ID)
long_data$Teacher <- as.factor(long_data$Teacher)
long_data$Grade <- as.factor(long_data$Grade)
long_data$Course <- as.factor(long_data$Course)
long_data$Period <- as.factor(long_data$Period)
long_data$Category <- as.factor(long_data$Category)
long_data$Last <- as.factor(long_data$Last)
long_data$First <- as.factor(long_data$First)
long_data$Gender <- as.factor(long_data$Gender)
View(long_data)

summary(long_data)
write_xlsx(long_data, "long_data.xlsx")
summary(long_data)

# Filter data for a specific category (e.g., Total)
total_data <- long_data %>% filter(Category == "Total")
View(total_data)
#write_xlsx(long_data, "survey long data stats.xlsx")

# Filter data for a specific category (e.g., Attitude)
att_data <- long_data %>% filter(Category == "Attitude")
View(att_data)
summary(att_data)
#write_xlsx(long_data, "survey long data stats.xlsx")

# Filter data for a specific category (e.g., Career)
career_data <- long_data %>% filter(Category == "Career")
View(career_data)
#write_xlsx(long_data, "survey long data stats.xlsx")

# Filter data for a specific category (e.g., Environment)
env_data <- long_data %>% filter(Category == "Environment")
View(env_data)
#write_xlsx(long_data, "survey long data stats.xlsx")

# Filter data for a specific category (e.g., Corals)
coral_data <- long_data %>% filter(Category == "Corals")
View(coral_data)
summary(coral_data)
#write_xlsx(long_data, "survey long data stats.xlsx")

# standardize scores
# Create a lookup table for max scores
max_scores_lookup <- data.frame(
  Category = c("Total", "Attitude", "Career", "Environment", "Corals"),
  MaxPoints = c(140, 40, 25, 40, 35)
)

# Add the max points to the dataset
long_data <- long_data %>%
  left_join(max_scores_lookup, by = "Category")

long_data <- long_data %>%
  mutate(StandardizedScore = (Score / MaxPoints) * 100)

head(long_data)
View(long_data)
```

# Check Assumptions
Category	Normality	Homogeneity	Recommended Test
Attitude	p > 0.05	p > 0.05	Repeated Measures ANOVA
Career	p > 0.05	p > 0.05	Repeated Measures ANOVA
Corals	p ≤ 0.05	p ≤ 0.05	Friedman Test (or Robust ANOVA if data permits)
Environment	p ≤ 0.05	p ≤ 0.05	Friedman Test (or Robust ANOVA if data permits)
Total	p ≤ 0.05	p ≤ 0.05	Friedman Test (or Robust ANOVA if data permits)
Friedmans test - needs balanced block design (not every student did every)
```{r}
# Check for missing values
sum(is.na(long_data))

# Check if the dataset is balanced
table(long_data$ID, long_data$Survey, long_data$Category)

# Perform Shapiro-Wilk test for each Survey and Category
normality_results <- long_data %>%
  group_by(Survey, Category) %>%
  summarise(p_value = shapiro.test(Score)$p.value)

print(normality_results)

# Initialize an empty data frame to store results
homogeneity_results <- data.frame(Category = character(), p_value = numeric(), stringsAsFactors = FALSE)

# Loop through each unique category
for (cat in unique(long_data$Category)) {
  # Subset the data for the current category
  category_data <- subset(long_data, Category == cat)
  
  # Run Levene's test
  levene_result <- leveneTest(Score ~ as.factor(Survey), data = category_data)
  
  # Extract p-value and store the result
  homogeneity_results <- rbind(homogeneity_results, 
                                data.frame(Category = cat, 
                                           p_value = levene_result$`Pr(>F)`[1]))
}

# Print the results
print(homogeneity_results)

# Example for one category (e.g., "Attitude")
attitude_data <- subset(long_data, Category == "Attitude")
leveneTest(Score ~ as.factor(Survey), data = attitude_data)

# Inspect variance
variance_check <- long_data %>%
  group_by(Survey, Category) %>%
  summarise(variance = var(Score, na.rm = TRUE))

print(variance_check)
```

# Total Score of Surveys
Using mixed effects model instead for unbalanced design and within and between variability 
```{r}
### Repeated Measures ANOVA
# Analyze each category separately
category_results <- list()

for (cat in unique(long_data$Category)) {
  # Filter data for the current category
  category_data <- subset(long_data, Category == cat)
  
  # Fit the repeated measures ANOVA
  category_results[[cat]] <- aov(Score ~ Survey + Error(ID / Survey), data = category_data)
  
  # Print results for the category
  cat("Results for Category:", cat, "\n")
  print(summary(category_results[[cat]]))
}

# Example for one category (e.g., "Attitude")
# total_data <- subset(long_data, Category == "Total")

# Fit a linear model
lm_total <- lm(Score ~ Survey, data = total_data)

# Perform pairwise comparisons for Survey levels
posthoc_total <- emmeans(lm_total, pairwise ~ Survey)

# Print pairwise comparison results
summary(posthoc_total)
# significant differences from survey 1 and 4 and 2 and 4 

### LME models
# Fit a linear mixed-effects model
lmer_model <- lmer(Score ~ Survey + (1 | ID), data = total_data)
summary(lmer_model)

# Post-hoc pairwise comparisons with emmeans
posthoc <- emmeans(lmer_model, pairwise ~ Survey)
summary(posthoc) # significant differences between 1-4, 2-4, 3-4, 1-3

# Fit the linear mixed-effects model
#lmer_model <- lmer(Score ~ Survey + Teacher + Period + Gender + Grade + (1 | ID), data = total_data)
lmer_model <- lmer(Score ~ Survey * Gender + Gender + Survey + Teacher + Period + Grade + (1 | ID), data = total_data)

# Model summary
summary(lmer_model)

# Test fixed effects
anova(lmer_model)

# Pairwise comparisons for Teacher
posthoc_teacher <- emmeans(lmer_model, pairwise ~ Teacher)
summary(posthoc_teacher) # not significant

# Pairwise comparisons for Course
posthoc_course <- emmeans(lmer_model, pairwise ~ Period)
summary(posthoc_course) # not significant

# Post-hoc comparisons for other predictors (e.g., Gender)
posthoc_gender <- emmeans(lmer_model, pairwise ~ Survey * Gender)
summary(posthoc_gender) # 1-4 and 2-4 Men and 1-4, 2-4 and 3-4 Femaile

# Post-hoc comparisons for other predictors (e.g., Gender)
posthoc_grade <- emmeans(lmer_model, pairwise ~ Grade)
summary(posthoc_grade) # not significant

# Post-hoc comparisons for other predictors (e.g., Gender)
posthoc_grade <- emmeans(lmer_model, pairwise ~ Gender)
summary(posthoc_grade) # not significant

# Post-hoc comparisons for other predictors (e.g., Gender)
posthoc_grade <- emmeans(lmer_model, pairwise ~ Survey)
summary(posthoc_grade) # 1-3, 1-4, 204 significant

# Plot Survey and Gender interaction effects
plot(emmeans(lmer_model, ~ Survey * Gender))

# Create a bar plot with error bars
total_summary <- total_data %>%
  group_by(Survey) %>%
  summarize(
    mean_score = mean(Score, na.rm = TRUE),
    se_score = sd(Score, na.rm = TRUE) / sqrt(n())
  )

total <- ggplot(total_summary, aes(x = Survey, y = mean_score)) +
  geom_bar(stat = "identity", fill = "#2C5C9B", alpha = 0.8) +
  geom_errorbar(aes(ymin = mean_score - se_score, ymax = mean_score + se_score), 
                width = 0.2, size = 1.5, color = "black") +
  labs(
       x = "Survey", 
       y = "Mean Total Score") +
  scale_x_discrete(labels = c("1" = "Control", "2" = "Traditional Curriculum", "3" = "Hybrid Learning", "4" = "Field Trip")) +
    geom_hline(yintercept = 140, linetype = "dashed", color = "gray", size = 1) +
  ylim(0, 140) +
  theme_minimal(base_size = 24)
total
save_plot("total.tiff", total, base_height = 12, base_width = 12)

# Summarize data by Survey and Gender
total_summary <- total_data %>%
  group_by(Survey, Gender) %>%
  summarize(
    mean_score = mean(Score, na.rm = TRUE),
    se_score = sd(Score, na.rm = TRUE) / sqrt(n()),
    .groups = "drop"
  )

# Plot interaction
custom_colors <- viridis(3, option = "magma")

total_gender <- ggplot(total_summary, aes(x = Survey, y = mean_score, fill = Gender)) +
  geom_bar(stat = "identity", position = position_dodge(), alpha = 0.8) +
  geom_errorbar(
    aes(ymin = mean_score - se_score, ymax = mean_score + se_score),
    position = position_dodge(width = 0.9),
    width = 0.2,
    color = "black"
  ) +
  scale_fill_manual(values = c("F" = "#F2B6F9", "M" = "#51B2BA")) +
  labs(
    x = "Survey",
    y = "Mean Total Score",
    fill = "Gender"
  ) +
  scale_x_discrete(labels = c("1" = "Control", "2" = "Traditional Curriculum", "3" = "Hybrid Learning", "4" = "Field Trip")) +
  theme_minimal(base_size = 24) +
  geom_hline(yintercept = 140, linetype = "dashed", color = "gray", size = 1) +
  ylim(0, 140)
total_gender
save_plot("total_gender.tiff", total_gender, base_height = 12, base_width = 14)

total.graphs <- total+total_gender
total.graphs
save_plot("total_graphs.tiff", total.graphs, base_height = 10, base_width = 28)
```

# Attitude 
```{r}
# Example for one category (e.g., "Attitude")
#attitude_data <- subset(long_data, Category == "Attitude")

# Fit a linear model
lm_attitude <- lm(Score ~ Survey, data = attitude_data)

# Perform pairwise comparisons for Survey levels
posthoc_attitude <- emmeans(lm_attitude, pairwise ~ Survey)

# Print pairwise comparison results
summary(posthoc_attitude) # survey 2-4 significant

### LME models
# Fit a linear mixed-effects model
lmer_model <- lmer(Score ~ Survey + (1 | ID), data = att_data)
summary(lmer_model)

# Post-hoc pairwise comparisons with emmeans
posthoc <- emmeans(lmer_model, pairwise ~ Survey)
summary(posthoc) # significant differences between 1-4, 2-4

# Fit the linear mixed-effects model
lmer_model <- lmer(Score ~ Survey * Gender + Survey + Gender + Teacher + Period + Grade + (1 | ID), data = att_data)

# Model summary
summary(lmer_model)

# Test fixed effects
anova(lmer_model)

# Pairwise comparisons for Teacher
posthoc_teacher <- emmeans(lmer_model, pairwise ~ Teacher)
summary(posthoc_teacher) # not significant

# Pairwise comparisons for Course
posthoc_course <- emmeans(lmer_model, pairwise ~ Period)
summary(posthoc_course) # not significant

# Post-hoc comparisons for other predictors (e.g., Gender)
posthoc_gender <- emmeans(lmer_model, pairwise ~ Survey * Gender)
summary(posthoc_gender) # survey 2-4 for men and 1-4 for women 

# Post-hoc comparisons for other predictors (e.g., Gender)
posthoc_grade <- emmeans(lmer_model, pairwise ~ Grade)
summary(posthoc_grade) # not significant

# Post-hoc comparisons for other predictors (e.g., Gender)
posthoc_grade <- emmeans(lmer_model, pairwise ~ Gender)
summary(posthoc_grade) # not significant

# Post-hoc comparisons for other predictors (e.g., Gender)
posthoc_grade <- emmeans(lmer_model, pairwise ~ Survey)
summary(posthoc_grade) # 1-4, 2-4 significant

# Plot Survey and Gender interaction effects
plot(emmeans(lmer_model, ~ Survey * Gender))

# Summarize mean and standard error for each survey
attitude_summary <- att_data %>%
  group_by(Survey) %>%
  summarize(
    mean_score = mean(Score, na.rm = TRUE),
    se_score = sd(Score, na.rm = TRUE) / sqrt(n())
  )

# Create a bar plot with error bars
att_total <- ggplot(attitude_summary, aes(x = Survey, y = mean_score)) +
  geom_bar(stat = "identity", fill = "#2C5C9B", alpha = 0.8) +
  geom_errorbar(aes(ymin = mean_score - se_score, ymax = mean_score + se_score), 
                width = 0.2, size = 1.5, color = "black") +
  labs(
       x = "Survey", 
       y = "Mean Attitude Score") +
  geom_hline(yintercept = 40, linetype = "dashed", color = "gray", size = 1) +
  scale_x_discrete(labels = c("1" = "Control", "2" = "Traditional Curriculum", "3" = "Hybrid Learning", "4" = "Field Trip")) +
  ylim(0,40) +
  theme_minimal(base_size = 24)
att_total
save_plot("att_total.tiff", att_total, base_height = 12, base_width = 12)

# Summarize data by Survey and Gender
att_summary <- att_data %>%
  group_by(Survey, Gender) %>%
  summarize(
    mean_score = mean(Score, na.rm = TRUE),
    se_score = sd(Score, na.rm = TRUE) / sqrt(n()),
    .groups = "drop"
  )

# Plot interaction
att_graph <- ggplot(att_summary, aes(x = Survey, y = mean_score, fill = Gender)) +
  geom_bar(stat = "identity", position = position_dodge(), alpha = 0.8) +
  geom_errorbar(
    aes(ymin = mean_score - se_score, ymax = mean_score + se_score),
    position = position_dodge(width = 0.9),
    width = 0.2,
    color = "black"
  ) +
  scale_fill_viridis(discrete = TRUE) + 
  scale_fill_manual(values = c("F" = "#F2B6F9", "M" = "#51B2BA")) +
  labs(
    x = "Survey",
    y = "Mean Attitude Score",
    fill = "Gender"
  ) +
  geom_hline(yintercept = 40, linetype = "dashed", color = "gray", size = 1) +
  scale_x_discrete(labels = c("1" = "Control", "2" = "Traditional Curriculum", "3" = "Hybrid Learning", "4" = "Field Trip")) +
  theme_minimal(base_size = 24) +
  ylim(0, 40)
att_graph
save_plot("att_graph.tiff", att_graph, base_height = 12, base_width = 14)
```
# Career 
```{r}
# Fit a linear model
lm_career <- lm(Score ~ Survey, data = career_data)

# Perform pairwise comparisons for Survey levels
posthoc_career <- emmeans(lm_career, pairwise ~ Survey)

# Print pairwise comparison results
summary(posthoc_career) # 1-4 and 2-4 significant 

### LME models
# Fit a linear mixed-effects model
lmer_model <- lmer(Score ~ Survey + (1 | ID), data = career_data)
summary(lmer_model)

# Post-hoc pairwise comparisons with emmeans
posthoc <- emmeans(lmer_model, pairwise ~ Survey)
summary(posthoc) # significant differences between 1-4, 2-4, and 3-4

# Fit the linear mixed-effects model
lmer_model <- lmer(Score ~ Survey * Gender + Survey + Gender + Teacher + Period + Grade + (1 | ID), data = career_data)

# Model summary
summary(lmer_model)

# Test fixed effects
anova(lmer_model)

# Pairwise comparisons for Teacher
posthoc_teacher <- emmeans(lmer_model, pairwise ~ Teacher)
summary(posthoc_teacher) # not significant

# Pairwise comparisons for Course
posthoc_course <- emmeans(lmer_model, pairwise ~ Period)
summary(posthoc_course) # not significant

# Post-hoc comparisons for other predictors (e.g., Gender)
posthoc_gender <- emmeans(lmer_model, pairwise ~ Survey * Gender)
summary(posthoc_gender) # 1-4 for women 

# Post-hoc comparisons for other predictors (e.g., Gender)
posthoc_grade <- emmeans(lmer_model, pairwise ~ Grade)
summary(posthoc_grade) # not significant

# Post-hoc comparisons for other predictors (e.g., Gender)
posthoc_grade <- emmeans(lmer_model, pairwise ~ Gender)
summary(posthoc_grade) # not significant

# Post-hoc comparisons for other predictors (e.g., Gender)
posthoc_grade <- emmeans(lmer_model, pairwise ~ Survey)
summary(posthoc_grade) # 1-4, 2-4 and 3-4 significant

# Summarize mean and standard error for each survey
career_summary <- career_data %>%
  group_by(Survey) %>%
  summarize(
    mean_score = mean(Score, na.rm = TRUE),
    se_score = sd(Score, na.rm = TRUE) / sqrt(n())
  )

# Create a bar plot with error bars
career_total <- ggplot(career_summary, aes(x = Survey, y = mean_score)) +
  geom_bar(stat = "identity", fill = "#2C5C9B", alpha = 0.8) +
  geom_errorbar(aes(ymin = mean_score - se_score, ymax = mean_score + se_score), 
                width = 0.2, size = 1.5, color = "black") +
  labs(
       x = "Survey", 
       y = "Mean Career Score") +
  geom_hline(yintercept = 25, linetype = "dashed", color = "gray", size = 1) +
  scale_x_discrete(labels = c("1" = "Control", "2" = "Traditional Curriculum", "3" = "Hybrid Learning", "4" = "Field Trip")) +
  ylim(0,25) +
  theme_minimal(base_size = 24)
career_total
save_plot("career_total.tiff", career_total, base_height = 12, base_width = 12)

# Plot Survey and Gender interaction effects
plot(emmeans(lmer_model, ~ Survey * Gender))

# Summarize data by Survey and Gender
career_summary <- career_data %>%
  group_by(Survey, Gender) %>%
  summarize(
    mean_score = mean(Score, na.rm = TRUE),
    se_score = sd(Score, na.rm = TRUE) / sqrt(n()),
    .groups = "drop"
  )

# Plot interaction
career_graph <- ggplot(career_summary, aes(x = Survey, y = mean_score, fill = Gender)) +
  geom_bar(stat = "identity", position = position_dodge(), alpha = 0.8) +
  geom_errorbar(
    aes(ymin = mean_score - se_score, ymax = mean_score + se_score),
    position = position_dodge(width = 0.9),
    width = 0.2,
    color = "black"
  ) +
  scale_fill_viridis(discrete = TRUE) + 
  scale_fill_manual(values = c("F" = "#F2B6F9", "M" = "#51B2BA")) +
  labs(
    x = "Survey",
    y = "Mean Career Score",
    fill = "Gender"
  ) +
  geom_hline(yintercept = 25, linetype = "dashed", color = "gray", size = 1) +
  scale_x_discrete(labels = c("1" = "Control", "2" = "Traditional Curriculum", "3" = "Hybrid Learning", "4" = "Field Trip")) +
  theme_minimal(base_size = 24) +
  ylim(0, 25)
career_graph
save_plot("career_graph.tiff", career_graph, base_height = 12, base_width = 14)
```

# Environment 
```{r}
# Fit a linear model
lm_env <- lm(Score ~ Survey, data = env_data)

# Perform pairwise comparisons for Survey levels
posthoc_env <- emmeans(lm_env, pairwise ~ Survey)

# Print pairwise comparison results
summary(posthoc_env) # no significance

### LME models
# Fit a linear mixed-effects model
lmer_model <- lmer(Score ~ Survey + (1 | ID), data = env_data)
summary(lmer_model)

# Post-hoc pairwise comparisons with emmeans
posthoc <- emmeans(lmer_model, pairwise ~ Survey)
summary(posthoc) # significant differences between 1-4, 2-4

# Fit the linear mixed-effects model
lmer_model <- lmer(Score ~ Survey * Gender + Survey + Gender + Teacher + Period + Grade + (1 | ID), data = env_data)

# Model summary
summary(lmer_model)

# Test fixed effects
anova(lmer_model)

# Pairwise comparisons for Teacher
posthoc_teacher <- emmeans(lmer_model, pairwise ~ Teacher)
summary(posthoc_teacher) # not significant

# Pairwise comparisons for Course
posthoc_course <- emmeans(lmer_model, pairwise ~ Period)
summary(posthoc_course) # not significant

# Post-hoc comparisons for other predictors (e.g., Gender)
posthoc_gender <- emmeans(lmer_model, pairwise ~ Survey * Gender)
summary(posthoc_gender) # 1-4 for women 

# Post-hoc comparisons for other predictors (e.g., Gender)
posthoc_grade <- emmeans(lmer_model, pairwise ~ Grade)
summary(posthoc_grade) # not significant

# Post-hoc comparisons for other predictors (e.g., Gender)
posthoc_grade <- emmeans(lmer_model, pairwise ~ Gender)
summary(posthoc_grade) # not significant

# Post-hoc comparisons for other predictors (e.g., Gender)
posthoc_grade <- emmeans(lmer_model, pairwise ~ Survey)
summary(posthoc_grade) # 1-4, 2-4 significant

# Summarize mean and standard error for each survey
env_summary <- env_data %>%
  group_by(Survey) %>%
  summarize(
    mean_score = mean(Score, na.rm = TRUE),
    se_score = sd(Score, na.rm = TRUE) / sqrt(n())
  )

# Create a bar plot with error bars
env_total <- ggplot(env_summary, aes(x = Survey, y = mean_score)) +
  geom_bar(stat = "identity", fill = "#2C5C9B", alpha = 0.8) +
  geom_errorbar(aes(ymin = mean_score - se_score, ymax = mean_score + se_score), 
                width = 0.2, size = 1.5, color = "black") +
  labs(
       x = "Survey", 
       y = "Mean Environment Score") +
  geom_hline(yintercept = 40, linetype = "dashed", color = "gray", size = 1) +
  scale_x_discrete(labels = c("1" = "Control", "2" = "Traditional Curriculum", "3" = "Hybrid Learning", "4" = "Field Trip")) +
  ylim(0,40) +
  theme_minimal(base_size = 24)
env_total
save_plot("env_total.tiff", env_total, base_height = 12, base_width = 12)

# Plot Survey and Gender interaction effects
plot(emmeans(lmer_model, ~ Survey * Gender))

# Summarize data by Survey and Gender
env_summary <- env_data %>%
  group_by(Survey, Gender) %>%
  summarize(
    mean_score = mean(Score, na.rm = TRUE),
    se_score = sd(Score, na.rm = TRUE) / sqrt(n()),
    .groups = "drop"
  )

# Plot interaction
env_graph <- ggplot(env_summary, aes(x = Survey, y = mean_score, fill = Gender)) +
  geom_bar(stat = "identity", position = position_dodge(), alpha = 0.8) +
  geom_errorbar(
    aes(ymin = mean_score - se_score, ymax = mean_score + se_score),
    position = position_dodge(width = 0.9),
    width = 0.2,
    color = "black"
  ) +
  scale_fill_viridis(discrete = TRUE) + 
  scale_fill_manual(values = c("F" = "#F2B6F9", "M" = "#51B2BA")) +
  labs(
    x = "Survey",
    y = "Mean Environment Score",
    fill = "Gender"
  ) +
  geom_hline(yintercept = 40, linetype = "dashed", color = "gray", size = 1) +
  scale_x_discrete(labels = c("1" = "Control", "2" = "Traditional Curriculum", "3" = "Hybrid Learning", "4" = "Field Trip")) +
  theme_minimal(base_size = 24) +
  ylim(0, 40)
env_graph
save_plot("env_graph.tiff", env_graph, base_height = 12, base_width = 14)
```
# Corals 
```{r}
# Fit a linear model
lm_coral <- lm(Score ~ Survey, data = coral_data)

# Perform pairwise comparisons for Survey levels
posthoc_coral <- emmeans(lm_coral, pairwise ~ Survey)

# Print pairwise comparison results
summary(posthoc_coral) # 1-3, 1-4, 2-4

### LME models
# Fit a linear mixed-effects model
lmer_model <- lmer(Score ~ Survey + (1 | ID), data = coral_data)
summary(lmer_model)

# Post-hoc pairwise comparisons with emmeans
posthoc <- emmeans(lmer_model, pairwise ~ Survey)
summary(posthoc) # significant differences between 1-3, 1-4, 2-4

# Fit the linear mixed-effects model
lmer_model <- lmer(Score ~ Survey * Gender + Survey + Gender + Teacher + Period + Grade + (1 | ID), data = coral_data)

# Model summary
summary(lmer_model)

# Test fixed effects
anova(lmer_model)

# Pairwise comparisons for Teacher
posthoc_teacher <- emmeans(lmer_model, pairwise ~ Teacher)
summary(posthoc_teacher) # not significant

# Pairwise comparisons for Course
posthoc_course <- emmeans(lmer_model, pairwise ~ Period)
summary(posthoc_course) # not significant

# Post-hoc comparisons for other predictors (e.g., Gender)
posthoc_gender <- emmeans(lmer_model, pairwise ~ Survey * Gender)
summary(posthoc_gender) # 1-3 and 1-4 and 2-3 and 2-4 for women and 1-4 for men

# Post-hoc comparisons for other predictors (e.g., Gender)
posthoc_grade <- emmeans(lmer_model, pairwise ~ Grade)
summary(posthoc_grade) # not significant

# Post-hoc comparisons for other predictors (e.g., Gender)
posthoc_grade <- emmeans(lmer_model, pairwise ~ Gender)
summary(posthoc_grade) # not significant

# Post-hoc comparisons for other predictors (e.g., Gender)
posthoc_grade <- emmeans(lmer_model, pairwise ~ Survey)
summary(posthoc_grade) # 1-3, 1-4, 2-4 significant

# Summarize mean and standard error for each survey
coral_summary <- coral_data %>%
  group_by(Survey) %>%
  summarize(
    mean_score = mean(Score, na.rm = TRUE),
    se_score = sd(Score, na.rm = TRUE) / sqrt(n())
  )

# Create a bar plot with error bars
coral_total <- ggplot(coral_summary, aes(x = Survey, y = mean_score)) +
  geom_bar(stat = "identity", fill = "#2C5C9B", alpha = 0.8) +
  geom_errorbar(aes(ymin = mean_score - se_score, ymax = mean_score + se_score), 
                width = 0.2, size = 1.5, color = "black") +
  labs(
       x = "Survey", 
       y = "Mean Coral Score") +
  geom_hline(yintercept = 35, linetype = "dashed", color = "gray", size = 1) +
  scale_x_discrete(labels = c("1" = "Control", "2" = "Traditional Curriculum", "3" = "Hybrid Learning", "4" = "Field Trip")) +
  ylim(0,35) +
  theme_minimal(base_size = 24)
coral_total
save_plot("coral_total.tiff", coral_total, base_height = 12, base_width = 12)

# Plot Survey and Gender interaction effects
plot(emmeans(lmer_model, ~ Survey * Gender))

# Summarize data by Survey and Gender
coral_summary <- coral_data %>%
  group_by(Survey, Gender) %>%
  summarize(
    mean_score = mean(Score, na.rm = TRUE),
    se_score = sd(Score, na.rm = TRUE) / sqrt(n()),
    .groups = "drop"
  )

# Plot interaction
coral_graph <- ggplot(coral_summary, aes(x = Survey, y = mean_score, fill = Gender)) +
  geom_bar(stat = "identity", position = position_dodge(), alpha = 0.8) +
  geom_errorbar(
    aes(ymin = mean_score - se_score, ymax = mean_score + se_score),
    position = position_dodge(width = 0.9),
    width = 0.2,
    color = "black"
  ) +
  scale_fill_viridis(discrete = TRUE) + 
  scale_fill_manual(values = c("F" = "#F2B6F9", "M" = "#51B2BA")) +
  labs(
    x = "Survey",
    y = "Mean Coral Score",
    fill = "Gender"
  ) +
  geom_hline(yintercept = 35, linetype = "dashed", color = "gray", size = 1) +
  scale_x_discrete(labels = c("1" = "Control", "2" = "Traditional Curriculum", "3" = "Hybrid Learning", "4" = "Field Trip")) +
  theme_minimal(base_size = 24) +
  ylim(0, 35)
coral_graph
save_plot("coral_graph.tiff", coral_graph, base_height = 12, base_width = 14)
```

# Gender 
```{r}
# final plot 
att_career <- att_total+att_graph+career_total+career_graph
plot_layout(ncol = 2, byrow = TRUE)
att_career
save_plot("att_career.tiff", att_career, base_height = 18, base_width = 26)

env_coral <- env_total+env_graph+coral_total+coral_graph
plot_layout(ncol = 2, byrow = TRUE)
env_coral
save_plot("env_coral.tiff", env_coral, base_height = 18, base_width = 26)

# Aggregate data for plotting
aggregated_data <- total_data %>%
  group_by(Survey, Gender) %>%
  summarise(MeanResponse = mean(Score), .groups = "drop")

# Create the bar graph
ggplot(aggregated_data, aes(x = factor(Survey), y = MeanResponse, fill = Gender)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  labs(
    title = "Survey Responses by Survey Number and Gender",
    x = "Survey Number",
    y = "Mean Response",
    fill = "Gender"
  ) +
  theme_minimal()

# Aggregate data for plotting
aggregated_data <- att_data %>%
  group_by(Survey, Gender) %>%
  summarise(MeanResponse = mean(Score), .groups = "drop")

# Create the bar graph
ggplot(aggregated_data, aes(x = factor(Survey), y = MeanResponse, fill = Gender)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  labs(
    title = "Survey Responses by Survey Number and Gender",
    x = "Survey Number",
    y = "Mean Response",
    fill = "Gender"
  ) +
  theme_minimal()

# Aggregate data for plotting
aggregated_data <- career_data %>%
  group_by(Survey, Gender) %>%
  summarise(MeanResponse = mean(Score), .groups = "drop")

# Create the bar graph
ggplot(aggregated_data, aes(x = factor(Survey), y = MeanResponse, fill = Gender)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  labs(
    title = "Survey Responses by Survey Number and Gender",
    x = "Survey Number",
    y = "Mean Response",
    fill = "Gender"
  ) +
  theme_minimal()

# Aggregate data for plotting
aggregated_data <- env_data %>%
  group_by(Survey, Gender) %>%
  summarise(MeanResponse = mean(Score), .groups = "drop")

# Create the bar graph
ggplot(aggregated_data, aes(x = factor(Survey), y = MeanResponse, fill = Gender)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  labs(
    title = "Survey Responses by Survey Number and Gender",
    x = "Survey Number",
    y = "Mean Response",
    fill = "Gender"
  ) +
  theme_minimal()

# Aggregate data for plotting
aggregated_data <- coral_data %>%
  group_by(Survey, Gender) %>%
  summarise(MeanResponse = mean(Score), .groups = "drop")
print(aggregated_data)
#na_rows <- coral_data[is.na(coral_data$Score), ]
#print(na_rows)

# Create the bar graph
ggplot(aggregated_data, aes(x = factor(Survey), y = MeanResponse, fill = Gender)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  labs(
    title = "Survey Responses by Survey Number and Gender",
    x = "Survey Number",
    y = "Mean Response",
    fill = "Gender"
  ) +
  theme_minimal()
```

# Interactions
```{r}
# Linear mixed-effects model with random effect for ID
model <- lmer(StandardizedScore ~ Gender * Survey * Category + (1 | ID), data = long_data)
model2 <- lmer(StandardizedScore ~ Survey * Category + (1 | ID), data = long_data)

# OR a simpler linear model if no random effects are needed
# model <- lm(Score ~ Gender * Survey * Category, data = total_data)

# Compute EMMs
em_results <- emmeans(model, ~ Gender * Survey | Category)
em_results2 <- emmeans(model2, ~ Survey | Category)

# Print the results
print(em_results)
print(em_results2)

# Pairwise comparisons for Gender and Survey within each Category
pairwise_results <- emmeans(model, pairwise ~ Gender * Survey | Category)
pairwise_results2 <- emmeans(model2, pairwise ~ Survey | Category)

# Print pairwise results
print(pairwise_results)
print(pairwise_results2)

# Summary of the model
summary(model)
anova(model)
summary(model2)
anova(model2)

# # Compute estimated marginal means for the interaction
# em_interaction <- emmeans(model, ~ Gender * Survey)
# 
# # View the results
# print(em_interaction)
# 
# # Perform pairwise comparisons for interaction terms
# pairwise_results <- emmeans(model, pairwise ~ Gender * Survey)
# print(pairwise_results)

# Compute EMMs for Gender and Survey by Category
em_results <- emmeans(model, ~ Gender * Survey | Category)
em_results2 <- emmeans(model2, ~ Survey | Category)

# Convert the EMMs to a data frame
em_results_df <- as.data.frame(em_results)
em_results_df2 <- as.data.frame(em_results2)

# Round numeric values for clarity
em_results_df <- em_results_df %>%
  mutate(
    EMM = round(emmean, 2),          # Estimated Marginal Mean
    SE = round(SE, 2),               # Standard Error
    `Lower CL` = round(lower.CL, 2), # Lower Confidence Limit
    `Upper CL` = round(upper.CL, 2)  # Upper Confidence Limit
  ) %>%
  select(Category, Gender, Survey, EMM, SE, `Lower CL`, `Upper CL`) # Select desired columns

em_results_df2 <- em_results_df2 %>%
  mutate(
    EMM = round(emmean, 2),          # Estimated Marginal Mean
    SE = round(SE, 2),               # Standard Error
    `Lower CL` = round(lower.CL, 2), # Lower Confidence Limit
    `Upper CL` = round(upper.CL, 2)  # Upper Confidence Limit
  ) %>%
  select(Category, Survey, EMM, SE, `Lower CL`, `Upper CL`) # Select desired column

# Create a formatted table
em_table <- em_results_df %>%
  gt() %>%
  tab_header(
    title = "Estimated Marginal Means by Gender and Survey for Each Category",
    subtitle = "Results from Linear Mixed-Effects Model"
  ) %>%
  cols_label(
    Category = "Category",
    Gender = "Gender",
    Survey = "Survey",
    EMM = "Estimated Marginal Mean",
    SE = "Standard Error",
    `Lower CL` = "Lower CI",
    `Upper CL` = "Upper CI"
  ) %>%
  fmt_number(columns = c(EMM, SE, `Lower CL`, `Upper CL`), decimals = 2)

em_table2 <- em_results_df2 %>%
  gt() %>%
  tab_header(
    title = "Estimated Marginal Means by Survey for Each Category",
    subtitle = "Results from Linear Mixed-Effects Model"
  ) %>%
  cols_label(
    Category = "Category",
    Survey = "Survey",
    EMM = "Estimated Marginal Mean",
    SE = "Standard Error",
    `Lower CL` = "Lower CI",
    `Upper CL` = "Upper CI"
  ) %>%
  fmt_number(columns = c(EMM, SE, `Lower CL`, `Upper CL`), decimals = 2)

# View the table
em_table
em_table2
```

# Calculate differences
for all students pooled and both genders corals were the greatest increase in score and environment was the lowest increase in score 
```{r}
# Calculate the mean score difference by Category
all_students_diff <- long_data %>%
  filter(Survey %in% c("1", "4")) %>%  # Filter only Survey 1 and 4
  group_by(Category, Survey) %>%
  summarize(mean_score = mean(StandardizedScore, na.rm = TRUE), .groups = "drop") %>%
  pivot_wider(names_from = Survey, values_from = mean_score, names_prefix = "Survey_") %>%
  mutate(diff = Survey_4 - Survey_1)  # Calculate difference

# Find the categories with the largest and smallest differences
max_increase <- all_students_diff %>% filter(diff == max(diff))
min_increase <- all_students_diff %>% filter(diff == min(diff))

# View results
print(all_students_diff)
print(max_increase)
print(min_increase)

# Calculate the mean score difference by Category and Gender
gender_diff <- long_data %>%
  filter(Survey %in% c("1", "4")) %>%
  group_by(Category, Gender, Survey) %>%
  summarize(mean_score = mean(StandardizedScore, na.rm = TRUE), .groups = "drop") %>%
  pivot_wider(names_from = Survey, values_from = mean_score, names_prefix = "Survey_") %>%
  mutate(diff = Survey_4 - Survey_1)  # Calculate difference

# Find the largest and smallest differences by Gender
max_increase_gender <- gender_diff %>% filter(diff == max(diff))
min_increase_gender <- gender_diff %>% filter(diff == min(diff))

# View results
print(gender_diff)
print(max_increase_gender)
print(min_increase_gender)
```

